\chapter[Mental Models of Password Managers]{Mental Models of Password Managers}\label{chap:mental_models_pwm}

An important spillover of our previous exploration is that password managers are more likely adopted the longer people had struggled with juggling passwords: Older participants in the third personality study were more likely to use a \gls{PWM}. We have already corroborated market surveys that indicate a generally low adoption rate of password management software. As discussed in Sections \ref{sec:rw:user-behavior} and \ref{sec:rw:pwm_generators}, it has been hypothesized that users do not fully trust third parties with their credentials, so there seems to be an urge to stay in charge. Consequently, most users still try to memorize their passwords. On the other hand, password managers do provide many usability and security benefits, but why do users fail to see them? To this end, we see a lack of understanding about how users make sense of password managers. Our goal was to understand users' mental models of password managers first and then identify opportunities to improve them, which could increase adoption rates. 
%RQs
We thus aimed to answer the following research questions: 1) How do users think a password manager works? 2) What are their reasons (not) to adopt one?

To answer these questions, Martin Prinz and I explored attitudes and understandings in semi-structured qualitative user interviews. To get a more complete picture, we interviewed both people who already use a \gls{PWM} and also people who prefer other coping strategies. The outcome of this investigation has been published as an extended abstract at SOUPS 2017 \cite{Prinz2017MentalModel}.

\section{Background and Context}
%%%
% technical side
%%%

% PWM have existed since xyz. -- not possible to point a specific origin. 
Password managers can be either built into web browsers or act as a standalone solution that is independent of the password's purpose. Dedicated password managers have existed since the mid to late 1990s. Web Confidential\footurl{http://www.web-confidential.com/}{16.02.2018} was probably one of the first programs to facilitate password management, when it first surfaced in 1998. However, which of the browsers was first to add password storage capabilities cannot be easily traced back, but all major browsers added this feature in the early 2000s. Given the long history of this kind of user aide, adoption is still at only 12\% \cite{Olmstead2017AmerciansCybersecurity}. Even security experts disagree on the specific security benefits of different implementations\footurl{https://www.wired.com/2015/07/websites-please-stop-blocking-password-managers-2015/}{16.02.2018}. If the auto-fill feature is enabled, this can be used to create digital footprints for individuals\footurl{https://freedom-to-tinker.com/2017/12/27/no-boundaries-for-user-identities-web-trackers-exploit-browser-login-managers/}{16.02.2018}

% different system architectures: built into browser, local software (dedicated programm and/or browser extension), distributed storage.
There are different architectures for handling passwords: offline password managers keep a database of encrypted passwords locally on the user's machine, while online managers provide more mobility because passwords are held on a server or a distributed storage solution \cite{McCarney2012Tapas}. \textit{KeePass} and \textit{Password Safe} are notable representatives for the offline storage paradigm, while the cloud-based approach is dominated by third-party solutions like \textit{LastPass}, \textit{1Password}, and \textit{Dashlane}. Browser vendors have also transitioned to store passwords in the cloud, e.g. Apple Keychain for Safari, or Google Smartlock for Chrome. 


% pwms are not universally accepted as the best solution.
% some websites block password managers (by disabling the ``paste'' event on their login form), because under some circumstances password managers can be used to track users. 


%%%
% user side:
%%%
% academic

% briefly summarize experiments focused around mental models of password managers
Don Norman chapter on mental models "Some observations on mental models": MM provide predictive and explanatory power for understanding an interaction \cite{Norman1982ObservationsMentalModels}

\cite{Kang2015MentalModelsDrawing} \cite{BravoLillo2011WarningsMentalModel}



\section{User Interviews}
Our main objective was to understand how users make sense of password support tools, in particular password managers. This would allow us to explore solutions that fit user expectations better. 

\subsection{Method and Protocol}
Focusing on attitudes and mental models requires a qualitative study method. We chose to conduct semi-structured, open-ended interviews. The sessions started with a thorough briefing about the study purpose and asking for permission to audio-record the conversation. The main questions were \textbf{(1)} \textit{Why do you need passwords?} and \textbf{(2)} \textit{What is your strategy to manage multiple accounts?}. From there, the study followed a more fine-grained follow-up questions to investigate the specifics of these two aspects. Moreover, Bravo-Lillo \etal showed the benefits of drawing tasks to find structural patterns in beliefs \cite{BravoLillo2011WarningsMentalModel}, so we also asked participants to \textbf{(3)} \textit{sketch how a password manager works}. This required that participants were aware of \glspl{PWM}. If they were not, they were told that it is a ``piece of software that stores a user's password'', which was expected to be vague enough to explore participants' expectations of this kind of software. The interviews took between five and 16 minutes. 

\subsubsection{Recruitment and Sample}
We first approached random passers-by on a popular street in Munich to obtain a diverse sample of participants. Six interviewees were recruited this way. However, these first six interviews showed that participants were unable to provide sufficient detail in answers to allow thorough analysis. Thus, we changed the recruiting method and approached employees of a design agency with which we had collaborated in the past. We also knew that the agency's policy required employees to utilize password managers. Moreover, the user group was more likely capable of visually expressing mental constructs, allowing for the envisioned analyses. Eight additional participants were interviewed in this sample, giving a total of N=14. They worked as experience designers and concept developers, and did not have formal training in computer science or engineering. The age of all interviewees ranged from 20 to 41. \todo{listen to recordings and determine gender distribution (CD in Munich)}. 

None of the interviewees from the street had used a password manager before, so we call this participant group the \textit{novices}. Since the \gls{PWM} was part of the company policy, all interviewees in the second group had used one before, so we call them the \textit{actives}. The separation allowed us to detect a shift of expectations before and after adopting a password manager. 

\subsubsection{Method Limitations}
The recruiting and sampling method is inherently limited. The novice user group was asked at a public spot, so it was difficult to provide enough context information and minimize distraction. On the plus side, we achieved diversity and the face-to-face set-up reassures trust, because it was clear that none of the information they gave us was going to be used to access their online accounts. \todo{XX} novices were unable to describe what a password manager was, before we gave them the above definition. Thus, their decision to refrain from using a \gls{PWM} was not made actively, but rather results from the lack of awareness. This limits the analysis of attitudes and self-reported behavior. Finally, the second user group has a homogeneous background: design and communication. All these limitations demand that the results not be generalized to a larger population. Instead, they should be seen as rough trends that help understand a first set of underlying mental models, rather than the entire spectrum thereof.

\subsection{Results}
% analysis approach

\subsubsection{Interviews}
\subsubsection{Drawing Tasks}


\section{Mental Model}


\section{Opportunities and Challenges}

\subsection{Customization and Personalization}

\subsection{Improving Sense of Agency}

\subsection{Leveraging Context}

\subsection{Limitations}

\section{Conclusion}


\noindent
\fbox{
	\hspace{1cm}
	\parbox[c][12cm]{0.7\linewidth}{
		\section*{Take Aways}
		\begin{itemize}[leftmargin=*]
			\item 
		\end{itemize}
	}
	\hspace{1cm}
}




