\subsection*{Acknowledgments}
Before I take you with me on a journey to the password universe, I want to say ``thank you'' to the people who have supported me throughout the years and who gave me the strength to finish this thesis. Words cannot describe how lucky I am to have (had) so many awesome colleagues, friends, and family members and I’m deeply humbled by your unconditional kindness. I would really love to give credit to each and everyone of you. But I need to embrace the essence of user experience and \textit{keep it simple}, so I’ll focus on the people without whom I simply could not have written this thesis. 

First and most of all, I want to thank my wife \textbf{Lena}, who couldn’t have known what she was getting into when we first met in Kindergarten. Thank you for constantly making me a better human and empowering me to pursue my dreams. I promise this was my first and last doctorate and I’ll have more time for you again.  

Working at the LMU media informatics group was incredibly great, thanks to so many fantastic people who all create an amazing work atmosphere. I have learned so much from my advisor \textbf{Heinrich Hußmann}, who always has an open ear for all sorts of affairs -- from bleeding edge web technology over research conundrums to personal matters. I wish I could deal with that amount of incoming requests as calmly as you. Thank you for enabling me to reach my goals! \textbf{Andreas Butz} and \textbf{Florian Alt} also have a great sense for hiring the right people. Thanks for your great input that sparked new ideas and corresponding publications. 

For the longest time, I shared an office with \textbf{Axel Hösl} and will really miss our daily philosophical conversations about punk and hip-hop subcultures, the broken system, the universe, and all the crazy things the Internet has to offer. I’m so glad that you live on my street, so I can just drop by to continue the conversation. Before Axel and I formed “437 Forschungsbande”, we were part of the “A206-crew” together with \textbf{Emanuel von Zezschwitz}. Ema has always been a special role-model because he is a workhorse who never complains about workload, he’s my favorite DJ, leader into debauchery, and ever helpful USEC rockstar -- I'm thankful for every bit of that. I deeply admire \textbf{Hanna Schneider} for her endless perseverance in so many situations. Thank you for opening my eyes for new lines of thoughts, your indulgence when I was being an idiot, and for empowering me to play Schafkopf. My work overlap with \textbf{Julie Wagner} lasted only a few months, but that was enough to become the best buddies. Thanks for sharing your passion for wild ideas, serendipity, and joint road-trips! \textbf{Sarah Tausch} is the most lovable person ever, a great neighbor who ensures our cats don’t starve, and a formidable jam partner. Thanks for your friendship and spending so much time with me. Thank you, \textbf{Henri Palleis} for sharing your dry humor and musical skills with me. Also, I hope one day you’ll let me buy one of your fantastic artworks. \textbf{Christian Mai} is one of a kind in terms of energy and a go-to guy to get a job done. Thanks for all the fun we had on our road trip through Colorado and for boosting my confidence all the time. \textbf{Daniel Buschek} sometimes intimidates me with his wisdom, but his matchless serenity make him an awesome person to hang out with in all circumstances. \textbf{Renate Häuslschmid} could snap my neck if she wanted to. Thank you for not doing that and being super nice and helpful instead. When \textbf{Malin Eiband} and I met for the first time, she thought I was an intern because I didn’t have a proper desk at the window at the time. Thanks for all the laughs we had over coffee / tea breaks in the kitchen. Thank you, \textbf{Mohamed Khamis}, for sharing your funky ideas -- and rooms at IDCs. When it was time to start writing the thesis, I was so glad to have \textbf{Nađa Terzimehic} to take over the majority of my teaching duties. Without her, I’d probably still torture students with the peculiarities of JavaScript. Finally, two people act behind the scenes, but they really are the backbone of the media informatics group: \textbf{Franziska Schwamb} and \textbf{Rainer Fink}. Franziska and I have always been in tune in terms of community spirit and helping each other out. Her comments made me laugh so often and without Franziska’s help I probably would have gone mad over expensing business travels. Thanks for everything, Franziska. I still don’t know how Rainer manages to do so many things at once. Whatever I asked for, I never expected to receive it right away, but Rainer delivers on requests faster than you can say “Windows-account termination policy enforcement”. All that without ever losing his unique sense of humor -- thank you!

Some people were actively involved in finalizing the thesis: I was smart enough to ask a bunch of folks to give feedback on the dissertation draft. \textbf{Hanna Schneider}, \textbf{Judith Kratzer}, \textbf{Julie Wagner}, \textbf{Lena Seitz}, \textbf{Malin Eiband}, \textbf{Mohamed Khamis}, \textbf{Molly Throgmorton}, \textbf{Sarah Tausch}, \textbf{Sophie Himmelstoß} - you are my heroes! Thanks for taking the time to provide your thoughts on such a dry topic. I still hope I can return this huge favor somehow. Speaking of favors: Thank you, \textbf{Alexander De Luca}, for making good things happen for me, although I’ve never really understood why you do this. Alex also connected me to \textbf{Kosta Beznosov} who volunteered to read the thesis, write a review, add tons of helpful suggestions, and travel to Munich to my defense -- thank you so much for the huge amount of time you sacrificed for me, although we had never met before. \textbf{Albrecht Schmidt} also jumped in as a committee member and helped close this chapter - thanks for your flexibility! I'll never forget the people with whom I started the PhD journey: \textbf{Marion Koelle} and \textbf{Patrick Lindemann} -- thank you for your great company and long-lasting friendship. 

Last, but not least, I want to express my gratitude to my family, but I prefer doing this in German. 
---
\selectlanguage{ngerman}
Danke \textbf{Irmgard} und \textbf{Klaus Stockinger}, dass ihr mir den Weg geebnet habt, um diesen Meilenstein zu erreichen. Danke für’s Zuhören wenn’s mal wieder nicht so voran ging wie gewollt und die Finanzspritzen, um mich mit Urlaub oder Musikkram abzulenken. Danke, dass ihr meine ab und an einsetzende Klugscheißerei überhört. Danke für euer Verständnis, wenn ich mal länger nicht angerufen habe oder wenn ich der Frage “und wie geht’s der Doktorarbeit?” ausgewichen bin. Danke, dass ihr mir immer zu verstehen gegeben habt, dass ihr glaubt ich schaffe das, und dass es auch okay wäre, wenn’s nicht klappt. Ist zum Glück alles gut gegangen. Ich möchte mich auch bei meinen Schwiegereltern \textbf{Claudia} und \textbf{Werner Seitz} bedanken, dass sie mich seit so langer Zeit bei sich zu Hause aufnehmen und versorgen. Ohne die Zeit im “Schreibexil” bei euch hätte sich der Abschluss der Arbeit noch viel weiter verzögert. Danke, dass ihr immer für mich da seid! \selectlanguage{english} ---

I wish I had space here to thank everybody who deserves to be mentioned. To deal with this limitation, I continue at \url{tobi-seitz.com/thanks}. Since I dedicated some research to emojis, the webpage lists a few more names and emojis that I associate with them. \linebreak
Thanks for being there! Let's roll!

\cleardoublepage