\section*{Preface}\label{preface}
This thesis, as so many other research projects, mostly tries to stay neutral and share insights into human behavior. Written in a scientific voice. The motivation to conduct research is often framed as mere problem solving. Before we reach that distance, let me tell you my personal history and explain why I picked this topic. 

I had tried to come up with passwords that were difficult to guess already at school. Explored different strategies (obfuscation, random strings). When I achieved to find a mnemonic device that allowed me to create pseudo-random passwords for hundreds of accounts, that felt awesome. I enjoyed when I succeeded to login to an account I hadn't used in years, with a password I could barely remember.  

However, I realized that not a lot of people do the same. But passwords take a notable part in people's digital lives. 

During the course of my PhD thesis, I talked to many people about my research questions. And in 99\% of the conversations, people could relate. Most of them shared their stories of their password agony and how they feel guilty because they have this and that strategy. People are interested in finding better solutions. In the past years I've noticed that more people ask me if we are still going to need passwords, because there are so many better solutions. Even non-security experts challenge your assumptions about security. Only few (I guess none) of them really shared my enthusiasm about creating memorable and strong passwords. 

However, after so many conversations had ended, people got back to me the next time I saw them and told me that they slightly changed their behavior or at least plan to do keep watching out more -- although I never told them to. This convinced me that it possible to do that on a larger scale. 

Although, as you'll find out, many of the projects to influence people's decisions and behavior show only small effect size, I still believe it is worthwhile to keep digging. For me personally, the thought of making a couple of billion people's lives even slightly better justified my spending years on the topic. 

%philantopist.