\chapter[Password Personality]{Understanding Password Selection Through the Lens of Personality Traits}\label{chap:pws_and_personality}

%\section{Introduction}
% general introduction
% show that not *everyone* does the same
% selection is an individual task, not everyone selects passwords in the same way
Although there are certain patterns, password selection is a task that each individual handles in their own way. 
% even perception differs.
We have shown in chapter \ref{chap:pasdjo} that password strength is even subjectively \textit{evaluated} differently depending on certain characteristics. Again, while there was an overall tendency, we were not able to break down the strength ratings by individual differences: was there a special user group who performed better than others in the game? What characterizes this user group? Moreover, our previous chapter shed light on password policies in the wild. In a few notable cases, the rules were challenging and reject a substantial part of passwords. Users have developed strategies to cope with rejected passwords, but it would be interesting to know the exact factors that contribute to their behavior in these circumstances. Egelman and Peer make a strong argument that there is no ``average user'' so it is necessary to look at individual differences to understand user behavior \cite{Egelman2015AverageUser}. 

% personality traits in cybersecurity are a thing
Demographic background is one of the external factors that influences password selection \cite{Mazurek2013Measuring,  Violettas2014PasswordsAvoidGreece, Wang2015ChinesePWs}. Personality traits have been brought into the discussion to explain user preferences, actions, and behaviors in security questions \cite{Brown2004GeneratingPWs, Gross2016EffectCognitiveEffort, Shropshire2006PersonalityITSec, Zakaria2013DesigningEffectiveSecurityMessages}. Especially in research about phishing susceptibility we find evidence that personality traits have the potential to explain behavior \cite{Halevi2013PilotStudyPersonality,Halevi2015SpearPhishing,ParrishJr2009PersonalityPhishing,Uebelacker2014SocialEngineering}. 
% personality traits even matter for passwords
Empirical results from password studies have been discussed and explained with different personality traits, too \cite{Haque2014PsychometricsStrongPassword,Weirich2001PrettyGoodPersuasion}. It is a evident that a user's personality shines through, when they select a word based on personal preferences. 
% characterize personality in this realm
Petrie classified users in distinct password personalities: family-oriented, fans, fantasists, and cryptics\footnote{The original survey is not available online anymore. In a personal inquiry with Ms Petrie, she said that the original data is with the firm who commissioned the survey. The aggregated statistics are available at \url{http://passwordresearch.com/stats/statistic130.html}, \la{29.01.2018}}. A LastPass report more roughly divides password usage into two groups \cite{LastPass2016PersonalitiesGetUsHacked}: Type A users want to stay in control and are driven to act securely, so they developed an elaborate system that they perceive as suitable. Type B users do not believe that their accounts are valuable to attackers, so they do not prioritize security over usability. 
% this kind of categorization is nice, but too rough, maybe something more sound would solve the situation.
Those two taxonomies stem from analysis of user-selected passwords, i.e. a retrospective evaluation. However, predictive approaches are under-explored. For instance, if a user is generally an emotional person, does this impact their password selection strategies? If a user is diligent in real life, do they invest effort to diligently craft passwords, too? 
%``I can't remember passwords anyway'' could be attributed to a person who is less conscientious and/or less neurotic. 
%Are neurotic people more concerned about password strength, because they fear attacks more?

% that's what we do.
In a series of user studies we explored the associations between personality traits and passwords. If such associations exist, a new range of support systems opens that are tailored to a user's personality. Current one-fits-all approaches could be re-designed radically. The research was carried out in cooperation with three students. In each separate project we focused on a different stage of the password life-cycle. Timo Erdelt investigated personality as predictor for the usability of composition policies \cite{Erdelt2017BA}. Paul Huber explored correlations between strength perceptions and personality traits \cite{Huber2016BA}. Finally, Aline Neumann examined personality factors in password selection \cite{Neumann2017BA}. In total, 453 individuals participated in three separate online studies. In the following chapter, the projects are put into context and their findings are discussed on a bigger picture. 

\subsubsection*{Research Objectives}
Our primary goal was to find ways to predict password behavior from personality traits. At this point, the discourse about risky passwords included personality factors as explanatory variables. However, only few empirical studies had been carried out to challenge the assumptions. We aimed to provide such empirical data and a discussion of the implications on the design of password policies and password authentication systems. For instance, adjusting requirements of password policies depending on the user's personality promises to reduce frustration of password selection. Thus, our over-arching research question can be framed as ``\textit{Does personality influence user's mental models of password strength and selection strategies?}''.

%more related work: Groß \etal \cite{Gross2016CognitiveDepletion}
%
%Goals/motivation: 
%- some publications tried interpreted their findings as a consequence of different personality traits. first mentioned in \cite{Weirich2001PrettyGoodPersuasion} that personality could make a difference -- however, this had never been tried to empirically measure. 
%- explore differences in certain coping strategies (selection and reuse) through the lens of personality
%- find suitable ways to adjust password policies \cite{Seitz2017PersonalizingPasswordPolicies} - when users need to reset their password, after they had used the service for a while. 
%
%We posed the following research questions before we set out to conduct the study.\\
%\textbf{RQ1 - Psychological Factors} How much do psychological factors affect the perceived strength of passwords?\\
%\textbf{RQ2 - Big-Five vs GDMS} Are the Big-Five traits stronger or weaker predictors for strength perception than other psychological variables?\\
%\textbf{RQ3 - Portfolio Factors} Is the personal password portfolio associated with strength perception?

\section{Background and Related Work}
%We position our work in usable security and privacy, in particular password research. Moreover, we include psychological models to better understand users dealing with passwords. 
In this section we give a brief overview about the characteristics of strong passwords and how users go about creating them. Moreover, we portray projects in usable security and privacy research in which the users' psyche has been the focus. 
% secondary
%premise
%influencing 
%preconditions


%Moreover, since sophisticated attacks often start with checking for already known passwords, obtaining clear text data goes along with attackers improving their approach. A password that was strong before can quickly become very weak.  
%TODO eventuell noch das DING WANG paper und oder das Neural Network Ding vom Melicher zitieren.

%\subsection{Studies of Personality in Cyber Security}
\subsection{Password Selection Preconditions and Context Factors}
%TODO subconscious? 
%not exhaustive
% DEMOGRAPHICS
However, apart from such conscious behavior, there may be other preconditions that make some users pick stronger passwords than others. In a large field study, Mazurek et al. found that computer science and engineering students created passwords that were less guessable than those from business or politics students \cite{Mazurek2013Measuring}. Beyond demographic background, context factors like the emotional state during password selection have also been investigated. 
% EMOTIONS
Gulenko examined the effect of presenting positive textual messages and icons during password selection and found benefits for the adoption of passphrases \cite{Gulenko2014PasswordsEmotion}. In contrast, putting users in a state of cognitive distress or depletion made participants choose weaker passwords in a large lab study \cite{Gross2016EffectCognitiveEffort}. 
% PAST EXPERIENCES / BREACHES / ATTACKS
Social pressure as another type of psychological leverage was investigated by Egelman \etal \cite{Egelman2013DoesMyPasswordGoUpToEleven}. While they argue that account value plays a superior role for the effectiveness of password meters, others have shown that the \textit{design} of a password meter does have a measurable impact on the effort users put into creating a password \cite{Ur2012HowDoesYourPasswordMeasureUp}. In summary, the literature shows that password selection depends on context factors beyond education and experience. 
\subsection{Personality Factors in Cyber Security}
In our work, we are interested in context factors of password strength originating from psychological variables like personality. One of the most commonly used models to characterize personality are the Big-Five traits (B5), also known as the five-factor model. Costa and McCrae \cite{Costa1992NEO} refer to the personality traits as \textit{openness to experience}, \textit{conscientiousness}, \textit{extraversion}, \textit{agreeableness}, and \textit{neuroticism} (OCEAN). The traits can be described with these exemplary adjectives \cite{McCrae1987ValidationFFM}:\\
\textbf{Openness:} imaginative, creative, curious, independent, liberal\\
\textbf{Conscientiousness:} careful, reliable, ambitious, scrupulous, neat, punctual\\
\textbf{Extraversion:} sociable, talkative, passionate, warm\\
\textbf{Agreeableness:} selfless, helpful, forgiving, cheerful, humble\\
\textbf{Neuroticism:} worrying, emotional, insecure, impatient, vulnerable, subjective

Most frequently, the influence of these personality traits have been explored for privacy-concerns, where the openness trait was associated with privacy attitudes \cite{Egelman2015PredictingAttitudes,Minkus2014PersonalizationPrivacy}. Other inquiries have shown that personality traits like neuroticism \cite{Halevi2013PilotStudyPersonality} or openness \cite{Uebelacker2014SocialEngineering} might be associated with the response to phishing attacks. The likelihood of employees adhering to security policies is potentially influenced by the manifestation of agreeableness and conscientiousness \cite{Shropshire2006PersonalityITSec,Shropshire2015}. These investigations show that personality trait models are a considerable factor in security and privacy. Yet, our understanding of the influence of personality on password perception and consequently password selection is still low. Our work tries to improve our understanding about the origin of the differences in users' judgments of password strength. 
%is linked to these aspects, because passwords protect privacy of users and sensitive data of companies. %However, how password behaviors are formed and how much of the  is explained by psychological factors is still underexplored

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%
%%%%
%%%%			STUDY  1 ONE EINS
%%%%			POLICIES
%%%%
%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{content/problem_space/pw-personality-study-1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%
%%%%
%%%%			STUDY  2 TWO ZWEI
%%%%			PERCEPTION
%%%%
%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{content/problem_space/pw-personality-study-2}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%
%%%%
%%%%			STUDY  3 THREE DREI
%%%%			SELECTION
%%%%
%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{content/problem_space/pw-personality-study-3}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%
%%%%
%%%%			GENERAL DISCUSSION
%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion of Research Questions}
In this section we interpret the results and hope to shed light on potential explanations for their origins. % and show the implications for password-authentication systems.

%TODO more elaborately discuss the origin of the results.

% Egelman says:
% Which psychographic segments should be targeted
% How can the segmentation be performed automatically without active user involvement (i.e.,
% How can knowledge of a user?s individual differences be abused, and how might this be prevented
% Which privacy and security mitigations are most likely to benefit from being tailored to individual traits

\subsection{Model selection}

%\subsection{RQ1: Traits of Password Strength Perception}
\subsection{The Weight of Openness and Conscientiousness}
%1. interpret negative correlation between openness and password rating
The passwords in our study received lower strength ratings from people with high openness scores, i.e. those who would describe themselves as creative, imaginative, intelligent or liberal. %At the same time, the meta passwords of these participants were similar to those of other participants. The effect thus only seems to concern the \textit{perception} of password strength -- not personal behavior. 
Looking for an explanation for these observations, we could attribute them to the ``divergent thinking'' facet exhibited by people with higher openness scores \cite{McCrae1987DivergentThinking} as well as the association with general intelligence \cite{DeYoung2015}. In this sense, participants with higher scores in openness could have been more likely to discover patterns in passwords due to higher creativity and cognitive abilities. A cue for this interpretation is P21's description of his approach: 
\begin{quote}
	\textit{``If it looked like dictionary passwords I rated it lower, more random unrepeated ones rated higher. Combinations of words also scored higher''}
\end{quote}
There were only few participants (6) who mentioned word combinations as beneficial for password strength. This diverges from the more commonly mentioned theme that dictionary words are bad for password strength. Accordingly, an explanation for the diverging assessment by participants with higher openness scores lies in the problem-solving skills that are usually attributed to more open people \cite{DeYoung2015, McCrae1987DivergentThinking}. The tasks we picked for our study are indeed problems that can be solved by relying on objective criteria. It is thus very likely that those participants took a more objective, albeit critical stance when rating the strength of the passwords. They correctly identified that password length is usually more important than the inclusion of digits. 

%2. interpret positive correlation with conscientiousness and LUDS
% policies have primed people too much and  
In contrast, conscientious participants were more likely to compare passwords by their variety of characters instead of length. It seems that they diligently check off an imaginary list of criteria that a strong password needs to have, just like when users face requirements dictated by policies.  Although much research has gone into optimizing password policies for a general audience, the recommendation of the 2word16 policy over more complex ones \cite{Shay2016DesigningPasswordPolicies}, for example, has not been implemented at scale in practice, yet. Instead, policies for the most part demand the inclusion of different character classes \cite{Wang2015EmperorsPolicies}. Our data suggest that conscientious participants, who are characterized as being more scrupulous and neat, are more likely to look out for details about character classes when they judge the strength of a password. Thus, we assume that conscientious participants benefit from having such a list of requirements dictated by policies and that they would not easily adopt policies demanding words instead of character diversity.
%The exposure to this type of policies might be the main reason why many participants in our study penalized passwords that contained words or that were in some ways meaningful.

We reason that managing passwords, i.e. \textit{creating}, \textit{committing to} and \textit{coping with} them \cite{Stobert2014PasswordLifeCycle}, is a problem that is more easily solvable for people with open minds or a disposition to treat the problem with diligence. It is less affected by their leaning towards agreeableness, extraversion, or neuroticism. So, we reason that password strength as an objective measure is also associated with the personality traits that are informed by objectivity. In that sense, the regression models understandably benefited from openness and conscientiousness.




\section{Implications for Research and Design} \label{sec:implications}
From the discussed points, we would like to point out where the findings have immediate and indirect implications. 

\subsection{Assessing Personality Traits in Password Studies}
Studying passwords is challenging, because we often lack access to real-world testbeds. Thus, online studies are the go-to method, which is usually reliable. However, the modification of independent variables in these online studies may not be the sole cause of effects. Confounding variables beyond demographic questions are seldom considered. Psychometrics may come in useful to evaluate causal relationships, however, they are to this date often neglected. We have shown that they have explanatory value and thus should not be neglected, if we want to stop designing security mitigations for the ``average user'' \cite{Egelman2015AverageUser}. 

However, including elaborate psychometric tests in addition to the core of the research question is probably unrealistic. Thus, to keep studies short and prevent participant fatigue, we propose to focus on the openness and conscientiousness personality traits. These can be measured with a bare minimum of 20 items, so they do not take more than five minutes to answer. If the scenarios in the password study resemble our comparison task, it is potentially viable to also include assessments on rational and intuitive decision making. 

Still, adding more items to a questionnaire might not always be possible. There is, however, a promising approach to quickly and inexpensively obtain personality data from study participants' past behavior. It is possible to infer personality facets from social media data, e.g. the number of friends on Facebook or public interactions \cite{Youyou2015Personality}. Youyou et al. found that such metrics can even outperform psychometric questionnaires. We thus suggest to request permission to access information that participants made available on such platforms and run personality analyses on the anonymized versions of it. Requesting permission to install software on mobile devices for research purposes has received little concern with crowd-workers in a large-scale study by Bentley and Chen \cite{Bentley2015Phonebook}. While we would not go so far as to read all contacts, call and message history, we believe a privacy sensitive version of this approach can replace personality tests.

\subsection{Password Suggestions}
Addressing personality facets might also become beneficial for real-time feedback during password creation. In the wild, we encounter password meters that estimate the strength of the password and in some cases even provide verbal feedback about what the user can do to improve the password (e.g. \cite{Wheeler2016zxcvbn}). A simple approach is to try and convince the user to pick a stronger password by suggesting a new one or a modified version of the entered password \cite{Forget2008ImprovingPasswordsThroughPersuasion, Seitz2016SuggestionsDecoy, Shay2015SpoonfulOfSugar}. 

Coming back to the finding that personality has an effect on how people \textit{compare} two passwords, we suppose that it plays a role for real-time suggestions, too. How a modified version or newly generated password is received likely depends on the user's personality. For instance, if the user's password is visible in clear text while they enter it, adding a suggestion next to the password results in a scenario similar to the comparison task from our study. Then we can conclude that it would be easier for conscientious people to assess the strength of their password, if there is a clear list of requirements than can be ``checked off'', to make it superior to the suggested password. Moreover, users with high openness trait characteristics might make their password longer if they see the suggested password is much longer than their own. The results from Seitz et al. point in this direction, because their study participants picked longer passwords, if a long password was suggested \cite{Seitz2016SuggestionsDecoy}. We reason that the approach could have produced larger effect sizes, had the participants' personality been taken into account and had the password been displayed in plain text. Similarly, user acceptance suffered in Shay et al.'s modification approach, maybe because it was not tailored to individual users \cite{Shay2015SpoonfulOfSugar}.

%\begin{itemize}
%\item there have been attempts to convince users of stronger passwords by simply suggesting new ones or alterations of the existing passwords \cite{Forget2008ImprovingPasswordsThroughPersuasion, Seitz2016SuggestionsDecoy, Shay2015SpoonfulOfSugar}. 
%\item If the user's password is visible during selection (a scenario most often useful on mobile devices \cite{Melicher2016UsabilityMobileTextPasswords}, this could make them compare the strength 
%\item there's a question if the users can stay neutral enough to re-evalute their own password 
%\end{itemize}


% DESIGN Space. Problem solving or not? 
\subsection{Personalized Password Policies}
Our results indicate that users have different, sometimes sophisticated approaches when they evaluate password strength, which is in line with related work \cite{Ur2016PerceptionsPassword}. The origin of these approaches is not directly linked to personality, but personality traits may influence what users pay attention to when they face password policies. Diligent users could look out for and remember a list of requirements, while only the minimum length is important for creative minds. Identity providers like Facebook or Google rely on passwords to authenticate users and currently implement a ``one-fits-all'' password policy. However, once users have used the services for a while, those parties already possess a large amount of user information. Consequently, the identity providers could alter the enforced policy for individual users depending on personality trait characteristics inferred from usage patterns. We propose to dynamically adjust the policy, for example when users reset their password. Conscientious users may welcome a more complex policy,  e.g. 3class12 (cf. Shay et al \cite{Shay2016DesigningPasswordPolicies}), while others create passwords with a 2word16 requirement. Such approaches could reduce user frustration and boost usability. 

%%TODO identity providers.
%\begin{itemize}
%\item our results indicate that users have different approaches to evaluating passwords
%\item since we see interactions with personality, we hypothesize that their personality influences what they
%take from password policies (e.g. what they can remember, or what they believe). 
%\item we cannot adjust / personalize password policies for unknown users
%\item we can personalize policies for password resets, because we already have some data about the user
%and it can be used to inform personality assessment.
%\item is the policy presented
%
%\end{itemize}


%Considering that people's personality is a factor that affects the perception of password strength, this directly leads to the question about how we can address personality in current password policies. The policies put a high burden on users and their effect is controversial. Thus, we could try and make password policies more usable by the framing them to better reach the right audiences in the right ways, rather than using ``one fits all'' approach. Of course, at the moment of password selection, a service does not know the user yet and thus lacks information on their personality. However, the problem might be solvable on mobile devices, where the usage patterns let us already draw conclusions about the user's personality \cite{Oliveira2011Psychographics}. The idea of adjusting password policies depending on the device on which the password is created has also been put forward recently \cite{Melicher2016UsabilityMobileTextPasswords}. It thus seems beneficial to investigate how to bring the two concepts together. 

\section{Conclusion and Future Work}
In our work we explored the relationship between personality traits, decision making styles and perceived password strength. An online survey with 100 participants revealed first indications of such relationships, most of which originated from the two personality traits openness and conscientiousness. We found that the Big-Five inventory is a moderate albeit non-negligible factor for a users' subjective perception of password strength. When passwords are compared, the rational and intuitive decision making style is also an important factor. 

\subsection{Contribution Summary}
% CONTRIBUTION STATEMENT:
Our work is one of the first to specifically investigate the relationship between personality traits and password strength, respectively password policies. We contribute evidence for (a) the existence of associations between the openness and consciousness traits and password strength assessment. We show that (b) users who show a rational decision making style perform worse in identifying stronger passwords than those who decide intuitively. Furthermore, we (c) conclude that the Big-Five traits were somewhat more predictive than general decision-making style, but that they should be combined depending on the problem. We also identified that (d) the composition of a personal password is also associated with the perception of other passwords. The contributions can help to inform the design of more usable password policies and real-time feedback, as outlined in Section \ref{sec:implications}.

\subsubsection{Future Work}
A possible next step is to try and leverage information about personality traits to tailor password feedback to make it more effective. As pointed out by related work, some users overestimate the strength improvement of using digits in passwords \cite{Ur2016PerceptionsPassword}, which we generally confirmed with our study. However, not all participants were equally influenced by digits or different character classes inside passwords. Respecting the personality type in real-time feedback while enforcing policies could make the messages more effective and boost usability at the same time. Beyond password policies, personality traits can inform choice architectures. For example, the Choose-Your-Own-Authentication approach \cite{Forget2015CYOA} can benefit from informed default settings depending on personality trait characteristics. The collection and aggregation of the necessary information could be done on mobile phones \cite{DeMontjoye2013}. We consider this approach promising because it has a good chance to improve the user experience of authentication mechanisms and it is already possible to achieve the goal with off-the-shelf technology.



\noindent
\fbox{
	\hspace{1cm}
	\parbox[c][8cm]{0.7\linewidth}{
		\section*{Take Aways}
		\begin{itemize}[leftmargin=*]
			\item Personality is weakly associated with all the measured dimensions: strength perception, policy preference / usability, and password selection. 
			\item The models for the perception of passphrases achieved the highest fit, suggesting a predictable association between personality and strength perception for this type of password. Comparing two passwords was associated with the conscientiousness traits. Mixed models that use both password features and personality trait scores as covariates are the most feasible approach.
			\item Older users might be the best target group for password support tools
			\item 
		\end{itemize}
	}
	\hspace{1cm}
}

