%!TEX root = ../../diss.tex

\chapter[Related Work Summary]{Related Work Summary}\label{chap:rw:summary}
%lingo: school of thought

\subsection*{Main Take-Aways}
% what is the current state of the world?
The current landscape of password research deals with both technical (Chapter \ref{chap:rw:passwords}) and human factors (Chapter \ref{chap:rw:user_perspective}). From a systems perspective, we can take away the following state of the world:
\begin{enumerate}
	\item There have been numerous attempts to replace passwords, but none of them are a panacea, because they entail a number of disadvantages that thwart their adoption. Any authentication scheme needs to withstand a plethora of attacks, but passwords seem to offer the best trade-off regarding usability, deployability and security compared to other schemes. Single-Sign On, multi-factor, as well as biometric authentication have gained importance and are likely to be further adopted in the future. Especially implicit biometric approaches can complement password-based authentication.
	\item Entropy as a password strength proxy only works for system generated passwords. Thus, the results of a number of studies would need to be reassessed or replicated through new studies. As of now, the most reliable metric is performing actual guessing attacks on passwords. Configuring attacks is not trivial but crucial for the results. Currently, the Password Guessing Services is seen as the most robust tool. If cracking is impossible (e.g. for proactive password checks), the best strength proxy is to estimate a guess number for skillful, informed, and resourceful attackers. Neural networks and zxcvbn are among the most useful proxies at this point, but it is worthwhile to triangulate to obtain the full picture.  
	\item Password strength is difficult to agree upon and highly context dependent. This makes it hard to give homogeneous, consistent advice to users.
\end{enumerate}
The technical aspects are contrasted and complemented by these human factors:
\begin{enumerate}
	\item Users need to manage many password-protected accounts, which is one of the drivers to come up with individual coping strategies. Although these strategies are often intuitively developed, they show commonalities among larger user groups. Passwords go through a life-cycle no matter what strategy is used. Much of user behavior can be classified as risky, but in many ways it is rational.
	\item Users tend to select predictable passwords, reuse them often, and write them down in physical or digital notes. While there are many hypotheses to explain user behavior, the problem is probably too nuanced to reach a final consensus regarding its origins. It appears to be an assessment on a per-user level.
	\item Many solutions have been proposed to guide and support users. Policies enforce requirements on users to mitigate strength issues. However, the design space of password policies is large and the best parameters are context dependent. Policies do not guarantee ``better'' passwords, but real time feedback can make them more bearable. Password managers can take some responsibility from users which had been shifted to them in previous years. Explicit user education was once seen as primary tool to combat risky behavior, but users continued to dismiss it, which has lead to the understanding that education alone does not solve all problems.
	\item Persuasive design has been a well-studied area in password authentication to aid user education. It leverages cognitive biases and heuristics to achieve more secure user behavior. Password meters encompass many persuasive strategies at once and offer a large design space to influence password selection. However, their real-world implementations have brought inconsistencies to light, which reduced their persuasive power. 
	\item Studying passwords has a multitude of facets and the method toolkit is large. Most studies rely on online-surveys (especially through \gls{mTurk}), usability tests in the lab, or qualitative interviews. Several principles for the design of valid studies in usable security and privacy have emerged in the past two decades. 
\end{enumerate}

\subsection*{Open Questions}
Based on the review of the related literature we have identified a number of questions that have not been answered to full extent. 
% what do we not already know about the current state of the world
% optimally, this motivates all the research questions for the following chapters?
\begin{enumerate}
	\item Since results regarding user behavior and knowledge about password strength are mixed, we still have not uncovered all context-factors and confounding variables for password selection. However, we need to have this kind of understanding if we aim to design better support systems to either reinforce or break certain habits. Since user education has been going on since decades, we ask: What have users learned about password security in the mean time? What are their mental models of password strength? Can we trace back certain behavior to failed mitigations? If personality traits are associated with privacy preferences, can we observe the same for password selection?
	
	\item Password reuse is rampant and still a great risk for users, primarily in the form of identity theft following social engineering attacks. However, this aspect has not been addressed by the design of password policies. If policies cannot always effectively combat weak passwords, do they prevent password reuse? Are there ways to detect password reuse at the opportune moment when users create an account? 
	
	\item 
\end{enumerate}

\subsection*{The Passionate Discourse About Passwords}\label{sec:rw:passionate_discourse}
As a meta summary, we observe that the often mixed or even contradictory results of empirical password studies have generated a very passionate discourse about them. 

Different schools of thought: one line of researchers want to influence user behavior (policies) and believes in the efficacy of making users pick stronger passwords and behave securely. Another group wants to avoid passwords as much as possible by advocating other authentication schemes. The third group is more fatalistic in the sense that they do not see much hope for security interventions because users are not going to react to them in the desired way (Herley, Florêncio)

.. perhaps the discourse might calm down somewhat if more replication studies were carried out. However, \textit{publication bias} plays a significant role in the decision to go through the effort of replicating a study. 


Summarize: what were the years when most work was done? what type of research was it? what are the most common methods?

It's a bit hard to pinpoint specific contributions, because the papers often shed light on certain nuances of password selection -- just like this thesis.

A lot of position papers, respectively theoretical papers with no empirical data.

CMU focuses on policies and tries tiny modifications of policies and feedback. It's like you almost can predict the next paper they are going to write. Stringent line of research (look at conclusions and wait for the paper).

%there are very few papers that try to replicate findings

\todo{restructure, this doesnt need to go here, merge with above section?}
talk about how critical and passionate this topic has been debated in the literature

--> surprisingly, a lot of position papers (sometimes pure argumentation, sometimes with mathematical/logical reasoning) 

Address Herley's, Florencio's, Sasse's work here. 

e.g. Counterfactuals, Bullying, Finite Effort 


central theme: solution XYZ is going to cure everything and will kill passwords in the long run. 


